### 2024-09-02 texts/paper_closing.tex
export with new numbers, rewritten sections (hypotheses, founder death)

% Created 2024-08-29 do 15:35
% Intended LaTeX compiler: pdflatex
\documentclass[12pt]{article}

@@

Given that operating at "the margin of financial sustainability" is the "ordinary condition" for cultural NPOs, \cite[p.2]{Licci_BaraldiBonini_2024_sustainability}, it is not surprising that costs have been listed as a primary reason for private museum closure:
\textcite{Adam_2021_rise} argues that "most of the cases of defunct private museums come back to the issue of cost" (p.79); a sentiment shared by \textcite{Walker_2019_collector} who notes that "private institutions struggle to contain escalating costs that are associated with running a museum" (p.234).
In other studies, "insufficient funding, high maintenance costs and lack of strong government support" \parencite[p.7]{Zennaro_2017_shanghai}. or a "shortage of fundings and interruptions in cash \parencite[p.45]{Song_2008_private}  were central reasons that forced private museums to close.
In a systematic review of closing reasons given in newspapers, magazines, social media accounts and personal exchange, \textcite{Velthuis_Gera_2024_fragility} find financial issues as the most commonly mentioned reasons for private museum closing.
Scholars have traced the closure of private museums to numerous types of declining financial situations, such as bankruptcies \parencite{Velthuis_Gera_2024_fragility,Liu_2019_identities,DeNigris_2018_museums}, costly legal battles \parencite{Velthuis_Gera_2024_fragility}, stock devaluations \parencite{Walker_2019_collector}, failing auction sales \parencite{Bechtler_Imhof_2018_future} or lack of government support \parencite{Zennaro_2017_shanghai,Velthuis_Gera_2024_fragility}.
\subsubsection*{diversification}

A central reason for financial vulnerability has been argued to be the crucial reliance on the single founder, who, as the iniators of their private museums, likely have contribute a large part of their institution's non-operating income \parencite{Frey_Meier_2002_beyeler}:
\textcite{Adam_2021_rise} argues that "ultimately, many of these failures demonstrate the fragility of spaces that rely on a single founder, whose motivations and financial stability may change, or who may not have realised the difficulties inherent in establishing their own art space" (p.82).
\textcite{Velthuis_Gera_2024_fragility} similarly argue that "because of their funding models and reliance on a sole founder, [private museums] are inherently fragile organizations" (p.1).
\textcite{Bechtler_Imhof_2018_future} also characterize "the dependency on single individuals only" as one of main reasons that "make these private museums fragile" (p.53).
\cite{StylianouLambert_etal_2014_museums} also characterizes many private institutions as "small individual museums that depend too much on their creators to guarantee their sustainability" (p.582).
Beyond a direct financial component, private museums have been also
characterized as fragile since they can only rely on their founder for "taste", "vision and drive" \parencite[p.77]{Adam_2021_rise}, "passion" \parencite[p.234]{Walker_2019_collector} or "skills and knowledge" \parencite[p.580]{StylianouLambert_etal_2014_museums}.
The link between diversification and mortality, i.e. higher survival chances for more diversified organizations, has been found for non-profit organizations more generally \parencite{Fernandez_2007_dissolution,Bielefeld_1994_survival,Hager_2001_vulnerability,Lu_Shon_Zhang_2019_dissolution}\footnote{An larger literature exists shows a link between diversification and measures of financial health and sustainability besides survival \parencite{Licci_BaraldiBonini_2024_sustainability}}; it is thus plausible to apply to museums as well.

@@

@@

\textbf{Hypothesis 2}: A founder's death increases the risk of closing.
\textbf{Hypothesis 2}: Reliance on the founder is the mechanism through which PMs are exposed to higher closing risk, thus founder-related variables can be used to predict closing.

@@

The mere availability of audiences and other potential donors might however not be sufficient if a museum is not perceived as deserving support; i.e. if its identity is found lacking. 
An identity that is legitimate (understandable) and favorable (positive) has been argued to have a vast number of positive outcomes for organizations \parencite{Lange_Lee_Dai_2010_reputation}.
As gaining the approval of different stakeholders is necessary for being understood and/or supported by other organizations in the environment, having a clearly defined and favorable identity influences resource acquisition and thereby survival prospects \parencite{Rao_1994_reputation}.
Even for non-profit organizations such as museums, which do not produce tangible goods (a domain where identity-related effects are especially prevalent, e.g. \cite{Hsu_2015_granted,Bogaert_etal_2014_ecological}), a clear and positive identity might be helpful in a number of ways:
In particular in the context of private museums, decline of legitimacy or reputation could lead to lower visitor numbers, less discounts from art dealers, less government grants or lower chances on collaborations with other museums or corporations, resulting in lower revenues and higher acquisition costs.

@@

% latex table generated in R 4.3.3 by xtable 1.8-4 package
% Wed Aug 28 11:41:43 2024
\begin{table}[ht]
\centering
\begin{tabular}{lrll}
  \hline 
 \multicolumn{1}{l}{Self-identification} & \multicolumn{1}{l}{N} & \multicolumn{1}{l}{Self-ID (recoded)} & \multicolumn{1}{l}{examples}\\ 
 \hline
 Museum & 190 & Museum & Museum Kampa, Museum Barberini \\ 
  Foundation & 104 & Foundation & Marciano Art Foundation, Hill Art Foundation \\ 
  Collection &  69 & Collection & The Farjam Collection, Sammlung Boros \\ 
  None &  64 & Other & The Bunker, The Broad \\ 
  Center &  17 & Other & Dairy Art Centre, Art Center Nabi \\ 
  Gallery &  14 & Other & Saatchi Gallery, Galerie C15 \\ 
  Art space &  13 & Other & El Espacio 23, Qiao Space \\ 
  House / villa &  11 & Other & Villa La Fleur, Casa Daros Rio \\ 
  Institute &  10 & Other & Instituto Inhotim, Woods Art Institute \\ 
  Kunsthalle &   8 & Other & Kunsthalle Würth, G2 Kunsthalle \\ 
  Park / garden &   3 & Other & Schlosspark Eyebesfeld, Il Giardino dei Lauri \\ 
   \hline
\end{tabular}
\caption{Selfidentification} 
\label{tbl:t_selfid}
\end{table}

@@

\bigbreak
\noindent
A number of control variables are used: 
On the level of the founder, gender constitutes a time-invariant variable; with possible values being male, female and couple.
Centrality of the founder's vision is measured via the proxy of whether the name of the founder is included in the name of the museum.

@@

% latex table generated in R 4.3.3 by xtable 1.8-4 package
% Wed Aug 28 11:41:41 2024
\begin{table}[ht]
\centering
\begin{tabular}{llrrrrrr}
  \hline
 & & \multicolumn{2}{c}{Museum} & \multicolumn{4}{c}{Museum-year} \\ 
\cmidrule(r){3-4}\cmidrule(r){5-8} \multicolumn{1}{l}{} & \multicolumn{1}{l}{Variable} & \multicolumn{1}{l}{Count} & \multicolumn{1}{l}{Mean} & \multicolumn{1}{l}{Mean} & \multicolumn{1}{l}{SD} & \multicolumn{1}{l}{Min.} & \multicolumn{1}{l}{Max.}\\ 
 \hline
  \multicolumn{8}{l}{\textbf{Founder}} \\ 
 & Gender - Male & 294 & 0.584 & 6.0e-01 &  0.49 & 0 & 1 \\ 
   & Gender - Female & 79 & 0.157 & 1.7e-01 &  0.37 & 0 & 1 \\ 
   & Gender - Couple & 130 & 0.258 & 2.3e-01 &  0.42 & 0 & 1 \\ 
   & Founder - alive &  &  & 9.0e-01 &  0.30 & 0 & 1 \\ 
   & Founder - died recently &  &  & 1.9e-02 &  0.13 & 0 & 1 \\ 
   & Founder - died 2+ years ago &  &  & 7.9e-02 &  0.27 & 0 & 1 \\ 
   \multicolumn{8}{l}{\textbf{Museum}} \\ 
 & Self-Identification - Museum & 190 & 0.378 & 4.1e-01 &  0.49 & 0 & 1 \\ 
   & Self-Identification - Foundation & 104 & 0.207 & 1.9e-01 &  0.39 & 0 & 1 \\ 
   & Self-Identification - Collection & 69 & 0.137 & 1.3e-01 &  0.34 & 0 & 1 \\ 
   & Self-Identification - Other & 140 & 0.278 & 2.7e-01 &  0.45 & 0 & 1 \\ 
   & Founder name in Museum name & 240 & 0.477 & 4.8e-01 &  0.50 & 0 & 1 \\ 
   & AN Ranking - Not Included &  &  & 8.0e-01 &  0.40 & 0 & 1 \\ 
   & AN Ranking - Included &  &  & 1.4e-01 &  0.35 & 0 & 1 \\ 
   & AN Ranking - Dropped &  &  & 5.3e-02 &  0.23 & 0 & 1 \\ 
   & Exhibition any &  &  & 6.3e-01 &  0.48 & 0 & 1 \\ 
   & Opening year &  &  & 2.0e+03 & 11.10 & 1960 & 2020 \\ 
   \multicolumn{8}{l}{\textbf{Environment}} \\ 
 & PM density (country) &  &  & 4.5e-01 &  1.26 & 0.00077 & 27.26 \\ 
   & Pop. (millions) within 10km &  &  & 1.3e+00 &  1.68 & 0.000087 & 10.81 \\ 
   & Nbr PM within 10km &  &  & 1.8e+00 &  3.14 & 0 & 16 \\ 
   & Region - Africa & 8 & 0.016 & 1.2e-02 &  0.11 & 0 & 1 \\ 
   & Region - Asia & 142 & 0.282 & 2.9e-01 &  0.45 & 0 & 1 \\ 
   & Region - Europe & 248 & 0.493 & 5.0e-01 &  0.50 & 0 & 1 \\ 
   & Region - Latin America & 23 & 0.046 & 4.2e-02 &  0.20 & 0 & 1 \\ 
   & Region - North America & 72 & 0.143 & 1.4e-01 &  0.34 & 0 & 1 \\ 
   & Region - Oceania & 10 & 0.020 & 1.9e-02 &  0.13 & 0 & 1 \\ 
   & Time Period 2000-2004 &  &  & 1.0e-01 &  0.30 & 0 & 1 \\ 
   & Time Period 2005-2009 &  &  & 1.7e-01 &  0.37 & 0 & 1 \\ 
   & Time Period 2010-2014 &  &  & 2.6e-01 &  0.44 & 0 & 1 \\ 
   & Time Period 2015-2019 &  &  & 3.3e-01 &  0.47 & 0 & 1 \\ 
   & Time Period 2020-2024 &  &  & 1.4e-01 &  0.35 & 0 & 1 \\ 
   & Great Recession (2008/09) &  &  & 7.6e-02 &  0.27 & 0 & 1 \\ 
   & Covid Pandemic (2020/21) &  &  & 1.4e-01 &  0.35 & 0 & 1 \\ 
   \hline
\end{tabular}
\caption{Summary Statistics} 
\label{tbl:t_sumstats}
\end{table}

@@

Next to average yearly closing rates, private museum closing can also investigated over its age (unfortunately, no comparison to previous studies can be done as age-dependent mortality estimates are not reported).
Figure \ref{fig:p_hazard} shows the hazard rate over time, figure \ref{fig:p_surv} the corresponding survival probability.
The hazard rate increases strongly from 0.10\% in the first two years, to 0.69\% in the first 5 years, to 1.20\% at year 8, around which it then fluctuates, leading to an average hazard of 1.16\% for the first 30 years. \footnote{While it is possible to calculate the average hazard over longer periods (e.g. the overall average hazard rate is 0.90\%), for ages above 30 only few observations are available, which make these predictions rather uncertain.}
The increase in the first years is likely due to what has been called "initial resource endowments" in organizational research \cite{Carroll_Khessina_2019_demography,Hannan_1998_mortality} (not to be confused with the practice of a particular endowment for e.g. a museum REF, i.e. a stock of money, often invested in stocks, whose capital gains are used for operations).
As the establishment of a museum requires considerable resources, their founders appear to allocate enough resources to keep the risk of closure small in the first years after opening.
Given that founders are likely aware that their endeavor is unlikely to generate a profit, they are not vulnerable to the liability of newness \parencite{Stinchcombe_1965_structure}.

Such analysis can be used to estimate the average life expectancy of a private museum.
While only a fraction of all opened museums has closed (68 out of 503), under the assumption of average closing risk of 0.73\% in the first 8 years and 1.20\% onwards, median life expectancy can be estimated as 52 years.
On first sight such findings seems to differ from a previous quantitative analysis of the database, which argued that "the median number of years that private museums were open before they closed is 10 years" \parencite[p.5]{Velthuis_Gera_2024_fragility}, as previously only closed museums were considered.
To contribute to this 
\subsubsection*{Regression Results}

% latex table generated in R 4.3.3 by xtable 1.8-4 package
% Wed Aug 28 11:41:53 2024
\begin{table}[ht]
\centering
\begin{tabular}{p{0mm}lD{)}{)}{8)3}}
  \hline 
 \multicolumn{1}{l}{} & \multicolumn{1}{l}{Variable} & \multicolumn{1}{l}{r\_pop4}\\ 
 \hline
  \multicolumn{3}{l}{\textbf{Founder}} \\ 
 & Gender - Female & -.01 \; (0.34) \\ 
   & Gender - Couple & -.11 \; (0.32) \\ 
   & Founder - died recently & 1.06 \; (0.61) \\ 
   & Founder - died 2+ years ago & -.09 \; (0.55) \\ 
   \multicolumn{3}{l}{\textbf{Museum}} \\ 
 & Self-Identification - Foundation & .21 \; (0.41) \\ 
   & Self-Identification - Collection & -.44 \; (0.54) \\ 
   & Self-Identification - Other & 1.09 \; (0.30)^{***} \\ 
   & AN Ranking - Included & .18 \; (0.34) \\ 
   & AN Ranking - Dropped & .47 \; (0.50) \\ 
   & Founder name in Museum name & .07 \; (0.28) \\ 
   \multicolumn{3}{l}{\textbf{Environment}} \\ 
 & PM density (country) & .62 \; (0.35) \\ 
   & PM density$^{2}$ (country) & -.03 \; (0.02) \\ 
   & Pop. (millions) within 10km & .27 \; (0.10)^{**} \\ 
   & Nbr PM within 10km & .27 \; (0.08)^{***} \\ 
   & Nbr PM (10km) * Pop (10km) & -.09 \; (0.03)^{**} \\ 
   & Great Recession (2008/09) & -1.59 \; (1.01) \\ 
   & Covid Pandemic (2020/21) & .40 \; (0.29) \\ 
   \hline
 & Museums & 503 \\ 
   & Museum-years & 6096 \\ 
   & Closures & 68 \\ 
   & log. Likelihood & -338.69 \\ 
   & AIC & 711.39 \\ 
   & BIC & 749.12 \\ 
   \hline 
 \multicolumn{3}{l}{\footnotesize{standard errors in parantheses.\textsuperscript{***}p $<$ 0.001;\textsuperscript{**}p $<$ 0.01;\textsuperscript{*}p $<$ 0.05.}}
\end{tabular}
\caption{Cox Proportional Hazards Regression Results} 
\label{tbl:t_reg_coxph}
\end{table}

Table \ref{tbl:t_reg_coxph} shows the results of Cox proportional hazard regression model.
Coefficients are interpeted as log hazard ratios, i.e. as multipliers of the baseline hazard rate.
\paragraph*{Control Variables}

Neither gender nor inclusion of founder name is significantly associated with mortality rates.
While generally organizations whose name include the name of their founder might be more centered around the founder, and hence display higher inertia, making them less able to adapt to changing environments, these results indicate that it does not seem to be case for private museums.
\paragraph*{Identity Variables}

Museums which include the term "museum" in their name are the least likely to close, while those with the term "foundation" or "collection" have higher closure rates (but not significantly higher).
Museums which include neither of these are the most likely to close, and have mortality rates more than three times as high (\(e\)\textsuperscript{1.18} = 3.25, p<0.01) than institutions including the term "museum".
The usage of unconventional names might reflect adherence to museum standards which makes museums more understandable and familiar to third parties (e.g. audiences, other museums, companies or the government) and thereby facilitates interactions which are beneficial for long-term survival (e.g. attracting visitors, cooperations, sponsorships or subsidies).
However, as founders can choose the name of their museum arbitrarily, they might choose a name without reference to organizational forms associated with perpetuity precisely because their intention is not an institution enduring in perpetuity but rather a more temporally bounded project.

Museums included in the Museum of the World Database are \(e\)\textsuperscript{-0.88} = 0.41 (p < 0.05) times as likely to close than museums which are not, yet this result might be produced by different mechanisms.
On the one hand, it might reflect that larger and richer institutions are both more likely to be noticed by the MOW and more likely to survive (MOW inclusion would then be a proxy for resources).
On the other hand, similar to unconventional naming patterns, lack of inclusion in the MOW could be an indicator of deviance from museum standards, which could increase the unfamiliarity of the museum vis-a-vis third parties, reducing the chance of survival-enhancing interactions.

The reputation of the founder is not statistically related to closure. 
Museums whose founders are not included in the ARTnews collector ranking (reference category) are the least likely to close, followed by museums whose founders are included (not statistically significant) and by founders who were included at some point but are no longer included (not statistically significant).
\paragraph*{Organizational transformation}

The death of the founder is not associated with a significant difference in private museum closure, i.e. museums are not significantly more likely to close after their founder has died.
This possibly indicates that private museums might be less dependent on the mood of a single person (or couple), and have in fact become institutionalized to an extent that heirs of founders feel as motivated to continue the original founders' efforts as these founders themselves.
\paragraph*{Competition}

>>>>

% Created 2024-08-31 za 13:35
% Intended LaTeX compiler: pdflatex
\documentclass[12pt]{article}

@@

Given that operating at "the margin of financial sustainability" is the "ordinary condition" for cultural NPOs, \cite[p.2]{Licci_BaraldiBonini_2024_sustainability}, it is not surprising that costs have been listed as a primary reason for private museum closure:
\textcite{Adam_2021_rise} argues that "most of the cases of defunct private museums come back to the issue of cost" (p.79); a sentiment shared by \textcite{Walker_2019_collector} who notes that "private institutions struggle to contain escalating costs that are associated with running a museum" (p.234).
In other studies, "insufficient funding, high maintenance costs and lack of strong government support" \parencite[p.7]{Zennaro_2017_shanghai}, a "shortage of fundings and interruptions in cash" \parencite[p.45]{Song_2008_private}, bankruptcies \parencite{Velthuis_Gera_2024_fragility,Liu_2019_identities,DeNigris_2018_museums}, costly legal battles \parencite{Velthuis_Gera_2024_fragility}, stock devaluations \parencite{Walker_2019_collector} or failing auction sales \parencite{Bechtler_Imhof_2018_future} were central reasons that forced private museums to close.
In a systematic review of closing reasons given in newspapers, magazines, social media accounts and personal exchanges, \textcite{Velthuis_Gera_2024_fragility} find financial issues as the most commonly mentioned reasons for private museum closing.
\subsubsection*{diversification}

A central reason for financial vulnerability has been argued to be the crucial reliance on the single founder, who, as the iniators of their private museums, likely have to contribute a large part of their institution's non-operating income \parencite{Frey_Meier_2002_beyeler}:
\textcite{Adam_2021_rise} argues that "ultimately, many of these failures demonstrate the fragility of spaces that rely on a single founder, whose motivations and financial stability may change, or who may not have realised the difficulties inherent in establishing their own art space" (p.82).
\textcite{Velthuis_Gera_2024_fragility} similarly argue that "because of their funding models and reliance on a sole founder, [private museums] are inherently fragile organizations" (p.1).
\textcite{Bechtler_Imhof_2018_future} also characterize "the dependency on single individuals only" as one of main reasons that "make these private museums fragile" (p.53).
\cite{StylianouLambert_etal_2014_museums} also characterizes many private institutions as "small individual museums that depend too much on their creators to guarantee their sustainability" (p.582).
Beyond a direct financial component, private museums have been also been characterized as fragile since they can only rely on their founder for "taste", "vision and drive" \parencite[p.77]{Adam_2021_rise}, "passion" \parencite[p.234]{Walker_2019_collector} or "skills and knowledge" \parencite[p.580]{StylianouLambert_etal_2014_museums}.
The link between diversification and mortality, i.e. higher survival chances for more diversified organizations, has been found for non-profit organizations more generally \parencite{Fernandez_2007_dissolution,Bielefeld_1994_survival,Hager_2001_vulnerability,Lu_Shon_Zhang_2019_dissolution}\footnote{An larger literature exists shows a link between diversification and measures of financial health and sustainability besides survival \parencite{Licci_BaraldiBonini_2024_sustainability}}; it is thus plausible to apply to museums as well.

@@

Overall, two related arguments emerge from the current literature:
First, individual private museums rely on heavily on their founders for survival; events that interrupt the founders ability to provide resources are therefore associated with a highler likelihood of closing.
Second, due to their low diversification (high reliance on the founder), private museums have are more fragile than museums where single individuals are not comparably central figures.

@@

\textbf{Hypothesis 2}: Reliance on the founder is the mechanism through which PMs are exposed to a higher closing risk, thus founder-related variables predict closing, in particular
\begin{itemize}
\item \textbf{Hypothesis 2a}: Private museums are more likely to close after the death of the founder.
\item \textbf{Hypothesis 2b}: Private museums are more likely to close after their founder is excluded from a prestigious ranking.
\item \textbf{Hypothesis 2c}: Private museums which carry their founder's name in the museum name are more likely to close.
\end{itemize}

@@

The mere availability of audiences and other potential donors might however not be sufficient if a museum is not perceived as deserving support; i.e. if its identity is found lacking. 
An identity that is legitimate (understandable) and favorable (positive) has been argued to have a vast number of positive outcomes for organizations \parencite{Lange_Lee_Dai_2010_reputation}.
As gaining the approval of different stakeholders is necessary for being understood and/or supported by other organizations in the environment, having a clearly defined and favorable identity influences resource acquisition and thereby survival prospects \parencite{Rao_1994_reputation}.
Even for non-profit organizations such as museums, which do not produce tangible goods (a domain where identity-related effects are especially prevalent, e.g. \cite{Hsu_2015_granted,Bogaert_etal_2014_ecological}), a clear and positive identity might be helpful in a number of ways:
In the context of private museums, decline of legitimacy or reputation could lead to lower visitor numbers, less discounts from art dealers, less government grants or lower chances on collaborations with other museums or corporations, resulting in lower revenues and higher acquisition costs.

@@

% latex table generated in R 4.3.3 by xtable 1.8-4 package
% Fri Aug 30 12:00:47 2024
\begin{table}[ht]
\centering
\begin{tabular}{lrll}
  \hline 
 \multicolumn{1}{l}{Self-identification} & \multicolumn{1}{l}{N} & \multicolumn{1}{l}{Self-ID (recoded)} & \multicolumn{1}{l}{examples}\\ 
 \hline
 Museum & 190 & Museum & Museum Kampa, Museum Barberini \\ 
  Foundation & 104 & Foundation & Marciano Art Foundation, Hill Art Foundation \\ 
  Collection &  69 & Collection & The Farjam Collection, Sammlung Boros \\ 
  None &  64 & Other & The Bunker, The Broad \\ 
  Center &  17 & Other & Dairy Art Centre, Art Center Nabi \\ 
  Gallery &  14 & Other & Saatchi Gallery, Galerie C15 \\ 
  Art space &  13 & Other & El Espacio 23, Qiao Space \\ 
  House / villa &  11 & Other & Villa La Fleur, Casa Daros Rio \\ 
  Institute &  10 & Other & Instituto Inhotim, Woods Art Institute \\ 
  Kunsthalle &   8 & Other & Kunsthalle Würth, G2 Kunsthalle \\ 
  Park / garden &   3 & Other & Schlosspark Eyebesfeld, Il Giardino dei Lauri \\ 
   \hline
\end{tabular}
\caption{Selfidentification} 
\label{tbl:t_selfid}
\end{table}

@@

\bigbreak
\noindent
A number of control variables are used: 
On the level of the founder, gender constitutes a time-invariant variable; with possible values being male, female and couple.
Women constitute a minority of the private museums founders, thus from a typicality perspective \parencite{Rosch_1975_family} museums by female founders would be expected to close more often as female founders would have a greater difficulties in securing resources third parties, presumably primarily governments and corporate sponsors.
However, women may also be particularly associated with philanthropic initiatives (unlike e.g. corporate activities, \cite{Milam_2013_artgirls}) to such an extent that devaluation of female initiatives may not necessarily be present.

Centrality of the founder's vision is measured via the proxy of whether the name of the founder is included in the name of the museum.

@@

% latex table generated in R 4.3.3 by xtable 1.8-4 package
% Fri Aug 30 12:00:46 2024
\begin{table}[ht]
\centering
\begin{tabular}{llrrrrrr}
  \hline
 & & \multicolumn{2}{c}{Museum} & \multicolumn{4}{c}{Museum-year} \\ 
\cmidrule(r){3-4}\cmidrule(r){5-8} \multicolumn{1}{l}{} & \multicolumn{1}{l}{Variable} & \multicolumn{1}{l}{Count} & \multicolumn{1}{l}{Mean} & \multicolumn{1}{l}{Mean} & \multicolumn{1}{l}{SD} & \multicolumn{1}{l}{Min.} & \multicolumn{1}{l}{Max.}\\ 
 \hline
  \multicolumn{8}{l}{\textbf{Founder}} \\ 
 & Gender - Male & 294 & 0.584 & 6.0e-01 &  0.49 & 0 & 1 \\ 
   & Gender - Female & 79 & 0.157 & 1.7e-01 &  0.37 & 0 & 1 \\ 
   & Gender - Couple & 130 & 0.258 & 2.3e-01 &  0.42 & 0 & 1 \\ 
   & Founder - alive &  &  & 9.0e-01 &  0.30 & 0 & 1 \\ 
   & Founder - died recently &  &  & 1.3e-02 &  0.11 & 0 & 1 \\ 
   & Founder - died 2+ years ago &  &  & 8.4e-02 &  0.28 & 0 & 1 \\ 
   & Founder died &  &  & 9.7e-02 &  0.30 & 0 & 1 \\ 
   \multicolumn{8}{l}{\textbf{Museum}} \\ 
 & Self-Identification - Museum & 190 & 0.378 & 4.1e-01 &  0.49 & 0 & 1 \\ 
   & Self-Identification - Foundation & 104 & 0.207 & 1.9e-01 &  0.39 & 0 & 1 \\ 
   & Self-Identification - Collection & 69 & 0.137 & 1.3e-01 &  0.34 & 0 & 1 \\ 
   & Self-Identification - Other & 140 & 0.278 & 2.7e-01 &  0.45 & 0 & 1 \\ 
   & Founder name in Museum name & 240 & 0.477 & 4.8e-01 &  0.50 & 0 & 1 \\ 
   & AN Ranking - Not Included &  &  & 8.0e-01 &  0.40 & 0 & 1 \\ 
   & AN Ranking - Included &  &  & 1.4e-01 &  0.35 & 0 & 1 \\ 
   & AN Ranking - Dropped &  &  & 5.3e-02 &  0.23 & 0 & 1 \\ 
   & Exhibition any &  &  & 6.3e-01 &  0.48 & 0 & 1 \\ 
   & Opening year &  &  & 2.0e+03 & 11.10 & 1960 & 2020 \\ 
   \multicolumn{8}{l}{\textbf{Environment}} \\ 
 & PM density (country) &  &  & 4.5e-01 &  1.26 & 0.00077 & 27.26 \\ 
   & Pop. (millions) within 10km &  &  & 1.3e+00 &  1.68 & 0.000087 & 10.81 \\ 
   & Nbr PM within 10km &  &  & 1.8e+00 &  3.14 & 0 & 16 \\ 
   & Region - Africa & 8 & 0.016 & 1.2e-02 &  0.11 & 0 & 1 \\ 
   & Region - Asia & 142 & 0.282 & 2.9e-01 &  0.45 & 0 & 1 \\ 
   & Region - Europe & 248 & 0.493 & 5.0e-01 &  0.50 & 0 & 1 \\ 
   & Region - Latin America & 23 & 0.046 & 4.2e-02 &  0.20 & 0 & 1 \\ 
   & Region - North America & 72 & 0.143 & 1.4e-01 &  0.34 & 0 & 1 \\ 
   & Region - Oceania & 10 & 0.020 & 1.9e-02 &  0.13 & 0 & 1 \\ 
   & Time Period 2000-2004 &  &  & 1.0e-01 &  0.30 & 0 & 1 \\ 
   & Time Period 2005-2009 &  &  & 1.7e-01 &  0.37 & 0 & 1 \\ 
   & Time Period 2010-2014 &  &  & 2.6e-01 &  0.44 & 0 & 1 \\ 
   & Time Period 2015-2019 &  &  & 3.3e-01 &  0.47 & 0 & 1 \\ 
   & Time Period 2020-2024 &  &  & 1.4e-01 &  0.35 & 0 & 1 \\ 
   & Great Recession (2008/09) &  &  & 7.6e-02 &  0.27 & 0 & 1 \\ 
   & Covid Pandemic (2020/21) &  &  & 1.4e-01 &  0.35 & 0 & 1 \\ 
   \hline
\end{tabular}
\caption{Summary Statistics} 
\label{tbl:t_sumstats}
\end{table}

@@

Next to average yearly closing rates, private museum closing can also be investigated over its age (unfortunately, no comparison to previous studies can be done as age-dependent mortality estimates are not reported).
Figure \ref{fig:p_hazard} shows the hazard rate over time, figure \ref{fig:p_surv} the corresponding survival probability.
The hazard rate increases strongly from 0.10\% in the first two years, to 0.69\% in the first 5 years, to 1.20\% at year 8, around which it then fluctuates, leading to an average hazard of 1.16\% for the first 30 years. \footnote{While it is possible to calculate the average hazard over longer periods (e.g. the overall average hazard rate is 0.90\%), for ages above 30 only few observations are available, which make these predictions rather uncertain.}

The increase in the first years is likely due to what has been called "initial resource endowments" in organizational research \cite{Carroll_Khessina_2019_demography,Hannan_1998_mortality} (not to be confused with the practice of a particular endowment for e.g. a museum REF, i.e. a stock of money, often invested in stocks, whose capital gains are used for operations).
As the establishment of a museum requires considerable resources, their founders appear to allocate enough resources to keep the risk of closure small in the first years after opening.
Given that founders are likely aware that their endeavor is unlikely to generate a profit, they are not vulnerable to the liability of newness \parencite{Stinchcombe_1965_structure}.

Such analysis can be used to estimate the average life expectancy of a private museum.
While only a fraction of all opened museums has closed (68 out of 503), under the assumption of average closing risk of 0.73\% in the first 8 years and 1.20\% onwards, median life expectancy can be estimated as approximately 52 years.
On first sight such findings seems to differ from a previous quantitative analysis of the database, which argued that "the median number of years that private museums were open before they closed is 10 years" \parencite[p.5]{Velthuis_Gera_2024_fragility}, as previously only closed museums were considered.
Focusing on the entire population of private museums instead, this analysis provides additional insights into the expected longevity of private art museums.
\subsubsection*{Regression Results}

% latex table generated in R 4.3.3 by xtable 1.8-4 package
% Fri Aug 30 12:00:45 2024
\begin{table}[ht]
\centering
\begin{tabular}{p{0mm}lD{)}{)}{8)3}}
  \hline 
 \multicolumn{1}{l}{} & \multicolumn{1}{l}{Variable} & \multicolumn{1}{l}{r\_pop4}\\ 
 \hline
  \multicolumn{3}{l}{\textbf{Founder}} \\ 
 & Gender - Female & -.04 \; (0.34) \\ 
   & Gender - Couple & -.14 \; (0.32) \\ 
   & Founder died & .28 \; (0.43) \\ 
   \multicolumn{3}{l}{\textbf{Museum}} \\ 
 & Self-Identification - Foundation & .20 \; (0.41) \\ 
   & Self-Identification - Collection & -.45 \; (0.54) \\ 
   & Self-Identification - Other & 1.09 \; (0.30)^{***} \\ 
   & AN Ranking - Included & .19 \; (0.34) \\ 
   & AN Ranking - Dropped & .46 \; (0.50) \\ 
   & Founder name in Museum name & .05 \; (0.28) \\ 
   \multicolumn{3}{l}{\textbf{Environment}} \\ 
 & PM density (country) & .65 \; (0.34) \\ 
   & PM density$^{2}$ (country) & -.03 \; (0.02) \\ 
   & Pop. (millions) within 10km & .27 \; (0.09)^{**} \\ 
   & Nbr PM within 10km & .26 \; (0.08)^{***} \\ 
   & Nbr PM (10km) * Pop (10km) & -.09 \; (0.03)^{**} \\ 
   & Great Recession (2008/09) & -1.59 \; (1.01) \\ 
   & Covid Pandemic (2020/21) & .42 \; (0.29) \\ 
   \hline
 & Museums & 503 \\ 
   & Museum-years & 6096 \\ 
   & Closures & 68 \\ 
   & log. Likelihood & -339.68 \\ 
   & AIC & 711.35 \\ 
   & BIC & 746.87 \\ 
   \hline 
 \multicolumn{3}{l}{\footnotesize{standard errors in parantheses.\textsuperscript{***}p $<$ 0.001;\textsuperscript{**}p $<$ 0.01;\textsuperscript{*}p $<$ 0.05.}}
\end{tabular}
\caption{Cox Proportional Hazards Regression Results} 
\label{tbl:t_reg_coxph}
\end{table}

To investigate which covariates are associated with increased risk of closing, table \ref{tbl:t_reg_coxph} shows the results of Cox proportional hazard regression model.
Coefficients are interpeted as log hazard ratios, i.e. as multipliers of the baseline hazard rate.
\paragraph*{Control Variables}

Neither gender nor external shocks (the Great Recession and the Covid-19 Pandemic) are associated with higher closing rates.

inclusion of founder name is significantly associated with mortality rates.
While generally organizations whose name include the name of their founder might be more centered around the founder, and hence display higher inertia, making them less able to adapt to changing environments, these results indicate that it does not seem to be case for private museums.
\paragraph*{Founder dependence}

% latex table generated in R 4.3.3 by xtable 1.8-4 package
% Fri Aug 30 12:08:04 2024
\begin{table}[ht]
\centering
\begin{tabular}{rD{)}{)}{8)3}D{)}{)}{8)3}}
  \hline 
 \multicolumn{1}{l}{Recent death length} & \multicolumn{1}{l}{Recently dead} & \multicolumn{1}{l}{Long dead}\\ 
 \hline
   1 & .97 \; (1.02) & .19 \; (0.46) \\ 
    2 & 1.37 \; (0.61)^{*} & -.16 \; (0.55) \\ 
    3 & 1.06 \; (0.61) & -.09 \; (0.55) \\ 
    4 & .83 \; (0.61) & -.02 \; (0.55) \\ 
    5 & .62 \; (0.61) & .07 \; (0.55) \\ 
    6 & .50 \; (0.61) & .13 \; (0.55) \\ 
    7 & .87 \; (0.49) & -.51 \; (0.75) \\ 
    8 & .80 \; (0.49) & -.45 \; (0.75) \\ 
    9 & .91 \; (0.46)^{*} & -1.11 \; (1.03) \\ 
   10 & .83 \; (0.46) & -1.03 \; (1.03) \\ 
   11 & .78 \; (0.46) & -.98 \; (1.03) \\ 
   12 & .71 \; (0.46) & -.90 \; (1.03) \\ 
   \hline
\end{tabular}
\caption{Cox PH regression results wiht different death configurations} 
\label{tbl:t_reg_coxph_deathcfg}
\end{table}

The death of the founder is not associated with a significant difference in private museum closure, i.e. museums are not significantly more likely to close after their founder has died.
It is in principle possible to construct an alternative measure of founder death by splitting the time after the death in a short-term and long-term period; the short-term period might then capture the disruption by the founder, whereas the long-term period would only be reached by museums who have successfully acquired alternative funding models and are hence relatively viable.
However, the choice of choosing a cut-off point between the short- and long-term periods is relatively arbitrary.
